The evolution of gene expression in the term placenta of viviparous mammals

### 2015-12-15 manuscript/manuscript.Rnw
discuss SSLP1

The top 50 non-housekeeping expressed genes as measured by \ac{FPKM}
in the term placenta of \textit{Canis familiaris} are shown in
\cref{tab:top50_dog_placenta} with a list of all genes in Supplemental
\cref{tab:all_canis_familiaris_placenta}. The gene with the highest
expression %
(\ac{FPKM}=\Sexpr{gene.expression("canis
  familiaris","ENSCAFG00000002115")}) %
is the hormone relaxin (\textit{RLN2}). Other highly expressed genes
among the top 50 (\cref{tab:top50_dog_placenta}) include genes
involved in the cytoskeleton (\textit{TMSB10}, \textit{TMSB4X}),
anticoagulation (\textit{TFPI2}, \textit{PLA2G7}), innate immunity
(\textit{LCN2}, \textit{GM2A}), lipid metabolism (\textit{APOE}),
inhibition of matrix metalloproteinases (\textit{TIMP2},
\textit{TIMP3}), and growth factors (\textit{PGF}, \textit{EGFR},
HTRA1). The gene \textit{LGALS1}, which shows evidence of
immunosupression during pregnancy
\cite{Than.ea2008:Emergenceofhormonaland}, shows high expression in
dog %
(\ac{FPKM}=\Sexpr{gene.expression("canis familiaris","LGALS1")}). %
In addition, the gene SSLP1, a gene expressed in the placentas of cow
and sheep but not present in primates (Ushizawa et al., 2009), is also
expressed at high levels
%
(\ac{FPKM}=\Sexpr{gene.expression("canis familiaris","SSLP1")}). %
There is no placental expression of the estrogen synthase gene
(CYP19A1) in dogs, however expression of HSD3B1 is relatively high
(\ac{FPKM}=\Sexpr{gene.expression("canis familiaris","HSD3B1")}).

@@

>>>>

The top 50 non-housekeeping expressed genes as measured by \ac{FPKM}
in the term placenta of \textit{Canis familiaris} are shown in
\cref{tab:top50_dog_placenta} with a list of all genes in Supplemental
\cref{tab:all_canis_familiaris_placenta}. The gene with the highest
expression %
(\ac{FPKM}=%
\Sexpr{gene.expression("canis familiaris","ENSCAFG00000002115")}) %
is the hormone relaxin (\textit{RLN2}). Other highly expressed genes
among the top 50 (\cref{tab:top50_dog_placenta}) include genes
involved in the cytoskeleton (\textit{TMSB10}, \textit{TMSB4X}),
anticoagulation (\textit{TFPI2}, \textit{PLA2G7}), innate immunity
(\textit{LCN2}, \textit{GM2A}), lipid metabolism (\textit{APOE}),
inhibition of matrix metalloproteinases (\textit{TIMP2},
\textit{TIMP3}), and growth factors (\textit{PGF}, \textit{EGFR},
\textit{HTRA1}). The gene \textit{LGALS1}, which shows evidence of
immunosupression during pregnancy
\cite{Than.ea2008:Emergenceofhormonaland}, is highly expressed %
(\ac{FPKM}=\Sexpr{gene.expression("canis familiaris","LGALS1")}). %
In addition, the gene \textit{SSLP1}, a gene expressed in the
placentas of \textit{Bos tarus} (%
\Sexpr{gene.expression("bos tarus","SSLP1")}, %
\cref{tab:top50_cow_placenta}) and \textit{Ovis Aries}
(\Sexpr{gene.expression("ovis aries","SSLP1")},
\cref{tab:top50_sheep_placenta}) but is not present in primates
\cite{Ushizawa.ea2009:Characterizationandexpressionanalysis}, is also
expressed at high levels
%
(\ac{FPKM}=\Sexpr{gene.expression("canis familiaris","SSLP1")}). %
There is no placental expression of the estrogen synthase gene
(CYP19A1) in dogs, however expression of HSD3B1 is relatively high
(\ac{FPKM}=\Sexpr{gene.expression("canis familiaris","HSD3B1")}).

@@

\begin{figure}
<<sslp1.evolution.tree,echo=FALSE,fig.width=12,fig.height=12,warning=FALSE,message=FALSE>>=

tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="SSLP1" & species=="homo sapiens",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),gene.tree.table,gene.tree.expression)
@ 
\caption{Expression of prolactin receptor}
\label{fig:prolactin_receptor_evolution_tree}
\end{figure}
